---
title: "Ternary plots in R"
author: "Mike Croucher"
date: "18 May 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation

We are going to make use of the [ggtern](http://www.ggtern.com/) package. Install this in R with the following command

```{r eval=FALSE}
install.packages('ggtern')
```

You should only ever need to do this once per machine you are using.

## Loading and checking the data

Ensure that the file **iii_new CSR_PV.csv** is in the current working directory and load it into the object `mydata` with the following command. Note that we are telling R that the row names are in column 1 of our csv file.

```{r}
mydata = read.csv('iii_new CSR_ALL.csv', row.names=1)
```

View the data by entering the object name

```{r}
mydata
```

We can look at any individual row, for example Q10, using this notation

```{r}
mydata["Q10",]
```

We need to ensure that each row sums to 1. Otherwise, a ternary plot is not suitable. Check that using the `rowSums` command.

```{r}
rowSums(mydata)
```

Some of the rows don't quite sum to 1. Probably a rounding issue of some sort. Perhaps worth checking though

## Making the ternary plot

Load `ggtern` and produce a basic plot. Save it as the file `plot1.png`

```{r warning=FALSE, message=FALSE}
library(ggtern)
```
```{r}
base1 = ggtern(data=mydata,aes(R,C,S)) + geom_point()
show(base1)
ggsave('plot1.png')
```

Plot the same data but with some more annotations. Save it as the file `plot2.png`:

```{r}
markers = c(rep('Killesberg',19),rep('Rathaus Complex',30))
base2 = ggtern(data=mydata,aes(R,C,S)) + geom_point(aes(shape=markers)) + theme_nomask()
show(base2)
ggsave('plot2.png')
```

Same as before but make the original points bigger. Save it as the file `plot3.png`

```{r}
markers = c(rep('Killesberg',19),rep('Rathaus Complex',30))
base3 = ggtern(data=mydata,aes(R,C,S)) + geom_point(aes(shape=markers)) + theme_nomask() +
        scale_shape_manual(values=c(16,17))
base3 = base2 + geom_point(data = mydata['ORIGINAL Rathaus',],size=5,shape=17)
base3 = base2 + geom_point(data = mydata['ORIGINAL Kberg',],size=5,shape=16)
show(base3)
ggsave('plot3.png')
```

Now, we plot just the killesberg data. Save it as the file `plot4.png`

```{r}
markers = c(rep('original',1),rep('North',9),rep('South',9))
killesberg = mydata[1:19,]
base4 = ggtern(data=killesberg,aes(R,C,S)) + 
        geom_point(aes(shape=markers))  + 
        scale_shape_manual(values=c(1,16,0)) +
        theme_nomask()
show(base4)
ggsave('plot4.png')
```

